### 2022-01-14 papers/01 Reply round 1/response_TSG.tex

\query[R3-6]{ Please, check Figure 7. The authors claim that storage help in shaving the peak of total power imported from the main grid, but this looks the same in both figures.  }
\ans{We thank the Reviewer for providing this observation. What we claim is that the storage shaves the peak of the sum of the total power imported from the main grid and that of locally generated power. Due to the choice of the cost parameters, which follows \cite{atzeni2013}, local generation costs are cheaper than importing cost. We have redone the simulations where we readjust the cost parameters and introduced a positive quadratic coefficient on $f_i^{\mathrm{di}}$. Based on this adjustment, we can now see in Figure 7 in the revised manuscript (as also shown below) that the storage units particularly shave the peak of total power imported from the main grid.}

\renewcommand{\thefigure}{7}
\begin{figure}[!h]
	\centering
		\includegraphics[width=0.6\linewidth]{figures/simB_0402_pt1_v3}
	\caption{	Incorporating storage units causes a peak-shaving effect on the sum of the total power imported from the main grid and the power locally generated.}
\end{figure}
\query[R3-7]{ As a conclusion, the paper presents an interesting mathematical approach for clearing P2P markets. On the another hand, some relevant topics in this area remain uncovered and highlighted as future work. All in all, in my opinion, is a very good work, and should be considered for publication if the issues raised during revision are handled. }
\ans{We thank the Reviewer for the very positive comment and recommendation.}

@@

\query[R4-3]{ In the introduction, the authors assert that they extend a preliminary work by including nonlinear operational constraints, which complicate the analysis. However, they consider a linear approximation of power-flow equations (equations (13)). In fact, it seems that the only nonlinear constraint in the model is (15a), although some quadratic terms appear in the objective function. Is there any justification for this? Probably, the statement must be relaxed.  }
\ans{We agree with the Reviewer that the statement should be made clearer. That statement concerns the analysis for the algorithm development as we now have a larger set of coupling constraints than that in \cite{belgioioso2020energy}. In order to have an efficient algorithm, we exploit the structure of these constraints in decomposing the problem, thus resulting in Algorithm 1 in the revised manuscript, which is more complicated than \cite[Algorithm 1]{belgioioso2020energy}. Based on this discussion, in the revised manuscript, we have modified the statement as follows:}

@@

{ \edit{``Our market formulation extends our preliminary work \cite{belgioioso2020energy} by including network operational constraints and system operators in the model, which complicate the analysis as we need to exploit the problem structure to derive an efficient algorithm.''}}

\query[R4-4]{ Related to the previous comment, one of the nonlinearities appears in the objective function related to dispatchable units (equation (2)). However, in the numerical studies the parameter $Q_{i}^{di}$ is taken as zero, becoming a linear objective function, and simplifying the model. The same happens with the objective function associated to storage units (equation (4)). Have you tested the effect of these parameters in the model? Further, if these parameters take a positive value, some of the conclusions presented in the numerical study could change. I suggest the authors to analyze this issue.  }
\ans{We thank the Reviewer for raising this point. For our proposed algorithm, we have a convergence guarantee for either linear or non-linear objective functions as long as they are locally convex, i.e., convex with respect to local decision variables. In other words, our proposed algorithm works with quadratic local cost functions. Nevertheless, in view of this comment, we have updated the simulations in Section IV-C, where we now use a quadratic cost of dispatchable production, i.e., $Q_{i}^{\mathrm{di}}>0$, for all $i \in \mc N^{\mathrm{di}}$, which varies from one unit to another. The results are shown in Figure {7}, where we still observe a peak-shaving effect caused by the introduction of storage units. As this manuscript focuses on the distributed clearing mechanism design, we leave the analysis of the effect of different possible objective functions on the clearing outcome as future work.}

@@

\query[R4-5]{The objective function associated to the power traded with the main grid seems not so natural. If the electricity unit price, $c_{h}^{mg}$ in equation (9) is considered linear (instead of quadratic) with respect to the total consumption, this would simplify this term, further becoming (11) a decomposable function. Could the authors elaborate what is the advantage of using a quadratic term here? Is worth the improvement it could add with respect to the difficulty it brings to the model?}
\ans{By the formulation in {(9)--(11)}, which is based on \cite{atzeni2013}, the electricity price from the main grid follows a linear inverse demand function where it increases linearly as the total consumption increases. On the other hand, if  $c^{\mathrm{mg}}$ in (9) is a linear function, the electricity unit price is assumed to be constant, which is not the case in the practice. In our view, the former represent better the electricity unit price than the latter, as supported by \cite{atzeni2013noncooperative} and some references therein. In addition, as discussed in \cite[Section II-D]{atzeni2013noncooperative}, the cost function in {(11)} can represent either the actual energy cost (as a result of energy generation and transmission costs by the main grid) or simply a pricing function designed to incentivize load-shifting by the end users.}

\query[R4-6]{ Taking into account comment 1, suppose that we consider a reformulation of the problem to avoid coupling objective functions, and we employ the Alternating Direction Methods of Multipliers (ADMM) to solve the P2P market game in (1). Could the authors elaborate if this would be possible? Would the application of the ADDM also guarantees the convergence to a variational GNE of the P2P market game? What are the advantages of using Algorithm 1 proposed in the paper to solve the problem?  }
\ans{{Indeed, the only trivial case in which v-GNEs and social optima, i.e., the minimizers of the social welfare, coincide is for games with decoupled objective functions. This fact can be shown, for example, by a direct comparison of their KKT systems. In this special case, the original v-GNE seeking problem and the social welfare minimization problem are equivalent. Therefore, the latter can be solved via existing distributed optimization algorithms, e.g. ADMM. Unfortunately, in general, it is not possible to trivially reformulate games with coupled objectives as optimization problems -- see R4-3 for a simple counter-example in which a straightforward decoupling of the local objective functions fails to preserve the original equilibria. This simple fact motivated many researchers in recent years to design ad-hoc algorithms for games and generalized games with coupled cost functions, see e.g.  \cite{yi2019,paccagnan2019,liang2017distributed, koshal2016distributed,gadjov2020single,belgioioso2020b}.
}
}

@@

\query[R4-7]{ Regarding to the scalability of the algorithms, the author evaluates the convergence of the algorithm with respect to the number of agents and the connectivity, but they do not provide any information about the computational times. This is relevant to evaluate if it could be applied in a real large-scale system. }
\ans{We thank the Reviewer for raising this point. We agree that the computational time is relevant in practice and thus we have added the remark below in Section IV-C. Please, refer to our answer to the comment [R1-7] for a detailed discussion on the applicability of the proposed market-clearing mechanism on real large-scale systems.} %\red{particularly for the simulations with different numbers of agents}

{\edit{`` The average computation times of the inner loops, i.e., Algorithm~2 and 3, obtained on a computer with Intel Xeon E5-2637 3.5 GHz processors and 128 GB of memory, are 74.5 ms and 1.13 s, respectively.
		''}}
%

 \query[R4-8]{ Given a prosumer $i$, when Algorithm 1 is applied, in order to update the dual variable related to reciprocity constraints, the value of the power traded with neighbors plays an important role. In detail, I mean $\{p^{tr}_{(j,i)}\}_{j\in \mc N_{i}}$, the forward trade estimate to prosumer $i$. It seems that, depending on the order the prosumers are considered, the coordination of these variables could change, so that the convergence of the algorithm. Hence, it could be interesting to analyze how this affect to the behavior of the algorithm. For example, could it benefit to consider first prosumers with less neighbors, since maybe they reach faster an equilibrium? }
 	\ans{{
In the proposed market-clearing mechanism, all the local (primal and dual) updates of the prosumers are performed in parallel. We explicitly clarified this point in the new algorithm layout (please see the revised Algorithm 1).
 	}}

@@

 \textit{Minor comments:}

 	\query[R4-9]{ The notation used to define the quadratic terms of the objective functions must be introduced previously as “Basic notation”.  }
 	\ans{We thank the Reviewer for this suggestion and we have introduced the notation as follows:}

 	\edit{{``For any $x \in \bb R^n$, $\|x\|_A^2 = x^\top A x,$ with square symmetric matrix $A \succ 0$.''}}

 	\query[R4-10]{ The references to the lines of Algorithm 1 (in page 7) appearing in the text must be revised. For example, in item (iv) of Remark 2 (page 6), it must appear “…..(lines: 28-31)”. In the box related to dual variable update of prosumer $i$, in Figure 2 (page 6), the line references must be: lines 19-24. }
 	\ans{We thank the Reviewer for noticing this errors. As we have updated the presentation of Algorithm 1 to improve readability, we have also modified Remark 2 accordingly.}

### 2022-01-14 papers/01 Reply round 1/response_TSG.tex

\query[R2-6]{Algorithm 1 looks unfriendly to readers. The explanation for the algorithm (section III-A) is also unclear. In this current version, readers may have no clue to follow the formulation/notations in the algorithm and Fig. 2. For example, what are the dual variables used for? What is the intuitive idea behind the algorithm? What are the step sizes used for? The authors are suggested to separate Algorithm 1 into several parts and explain each part. }
\ans{ We agree with the Reviewer. Thus, following their suggestion, we have split the main algorithm in 3 parts: (1) (the revised) Algorithm 1, i.e., the main semi-decentralized loop, in which we summarize the information flow between prosumers and DNO, including the precise order of their updates; (2) Algorithm 2, i.e., the local primal-dual updates of the prosumers; (3) Algorithm 3, i.e., the central primal-dual update of the DNO.
%
Moreover, we have included a new paragraph in Section III-B to list all primal, dual, and auxiliary variables of (the revised) Algorithm 1, along with an explanation of their scopes within the algorithm and their economical and system-theoretic interpretations.
%
Finally, we have also re-designed the flow chart illustration of Algorithm 1 (Fig. 2).
%
Due to space limitations, we have omitted a more detailed discussion of the algorithm derivation. We refer the interested reader to \cite{belgioioso2020semi} for a complete discussion on this matter, including the analysis of the feasible step-sizes.
}

\newpage
\section*{Answers to Reviewer 3 }
\textit{This paper presents an interesting work dealing with the clearing of P2P energy markets while also considering technical constraints of the distribution network. From my point of view, some issues need to be further clarified:}

\query[R3-1]{Papers dealing with P2P markets are quite numerous recently. Literature review section is well structured and provides a fair number of works in this area, however the research gaps are not clearly stated. As a consequence, the contributions of the paper are quite generic. Please, consider to state, in a more specific way, the contributions of the paper. In particular, why the GNE approach is advantageous with respect to other decentralized frameworks proposed in the literature.  }
\ans{
{
In the revised introduction, we have
% highlighted the research gaps of the literature and rephrased 
improved the paragraphs on contribution to promote clarity.
The main advantage of game-theoretic energy market models with respect to optimization-based ones is their ability to formulate strategic (or self-interested) objectives, that naturally arise since prosumers are first and foremost self-interested decision-makers. Among game-theoretic models, GNE-based ones can additionally encode the global limitations of the grid (e.g., power-flow constraints, etc.) within the market.
%
Therefore, solutions of the resulting market clearing problem (i.e., variational GNEs) are economically efficient (locally and globally), strategically stable (no prosumer has an incentive in deviating from its current strategy), operationally safe (the global limitations of the grid are respected), and socially fair (the marginal loss for satisfying the grid constraints is the same for each prosumer). 
%
Unfortunately, the design of market-clearing mechanisms for game-theoretic GNE-based market models was (until very recently) elusive, due to the huge computation and communication overhead that, in practice, makes these methods inadequate for implementation on real large-scale systems.
%
In this paper, we exploit some recent advances in algorithmic game theory for generalized games \cite{belgioioso2020semi} to design a decentralized and scalable market-clearing algorithm, and show, via numerical simulations, that it is suitable for implementation in real large-scale systems.
}
}

@@

\newpage
\section*{Answers to Reviewer 4}
\textit{This paper presents a mathematical model for a P2P energy market formulated as a generalized game. They introduce in the game a new agent, a Distribution Network Operator (DNO), who is responsible of ensure the operational constraints of the market. Further, they propose a semi-decentralized algorithm to solve the problem, ensuring convergence to a Nash equilibrium. Finally, a numerical study is used to show the relevance of the market framework and the scalability of the algorithm.}

\query[R4-1]{In my opinion, the paper is clearly written, with a detailed explanation of the mathematical aspects, and introduces some relevant results. However, some questions need to be addressed in order to improve the quality of the work.  }
\ans{We thank the Reviewer for the positive comment.}

\textit{Comments and suggestions: }

\query[R4-2]{The authors state that due to the presence of coupling objective and constraints in the formulation, this implies the inapplicability of other decentralized algorithms employed in the literature (lines 54-57, page 2). This is not strictly true, since it is straightforward to reformulate the problem to decompose the objective function, just by replicating the variables avoiding the decomposition, and introduce new complicating constraints to coordinate these variables. Reformulate this sentence. }
\ans{{
What we meant is that games with coupling constraints and coupling objectives cannot, in general, be recast as distributed optimization problems, and therefore be solved via decentralized optimization algorithms.
%
In fact, the original market game (1) and the ``decomposed'' one, obtained by including local slack variables and coupling constraints to coordinate these variables have different equilibria -- consider the simple example at the end of this answer. Therefore, social-welfare optimization methods (e.g., ADMM) cannot be straightforwardly used to solve our game-theoretic problem. This is not surprising since variational GNEs and social optima, i.e., the minimizers of the social welfare, are essentially different configurations and coincide only in special cases. To avoid confusion, in the revised paper, we removed the claim of inapplicability.}}

{
\textit{Example (\textit{Decomposed games do not preserve the original equilibria}).}\\
Consider the following 2-player aggregative game:
\begin{align*} 
\text{player }1:& \quad \min_{u_1 \in \mathbb R}J_1(u_1,u_2) = 2(u_1+u_2 + 1)^\top u_1,\\
\text{player }2:& \quad \min_{u_2 \in \mathbb R} J_2(u_2,u_1) = (u_1+u_2 + 1)^\top u_2.
\end{align*}
Its Nash equilibria can be found in closed form by solving the system of equations obtained by imposing the pseudo-gradient mapping of the game to zero, i.e.,
\begin{equation*}
\left[
\begin{array}{l}
\nabla_{u_1} J_1(u_1,u_2)\\
\nabla_{u_2} J_2(u_1,u_2)
\end{array}
\right]
=
\left[
\begin{array}{l}
4u_1+2u_2+2\\
u_1+2u_2+1
\end{array}
\right]
\overset{!}{=}
\left[
\begin{array}{l}
0\\
0
\end{array}
\right],
\end{equation*} 
which yields a unique NE solution: $(u_1^*,u_2^*) = (-1/3,-1/3) $.\\
Now, consider instead the generalized game obtained by decomposing the objective functions using the local slack variables $\sigma_1,\sigma_2$ and the extra coupling constraints $\sigma_1 - u_2 = 0, \, \sigma_2 - u_1$:
\begin{align*} \
\text{player }1:& \quad 
\left\{
\begin{array}{r l}
\displaystyle
\min_{u_1,\sigma_1 \in \mathbb R}& J_1(u_1,\sigma_1) = 2(u_1+\sigma_1 + 1)^\top u_1,\\
\text{s.t. } & \sigma_1 - u_2 = 0\\
            &  \sigma_2 - u_1 = 0
\end{array}
\right.
\\
\text{player }2:& \quad 
\left\{
\begin{array}{r l}
\displaystyle
\min_{u_2,\sigma_2 \in \mathbb R}& J_2(u_2,\sigma_2) = (\sigma_2+u_2 + 1)^\top u_2,\\
\text{s.t. } & \sigma_1 - u_2 = 0\\
            &  \sigma_2 - u_1 = 0
\end{array}
\right.
\end{align*} 
Let us define the Lagrangian functions
\begin{align*}
\text{player }1: \quad \mc L_1(u_1,u_2,\sigma_1,\sigma_2,\lambda_{1,1},\lambda_{1,2}) &= 2(u_1+\sigma_1 + 1)^\top u_1 + \lambda_{1,1}^\top(\sigma_1 - u_2) + \lambda_{1,2}^\top(\sigma_2 - u_1),\\
\text{player }2: \quad \mc L_2(u_2,u_1,\sigma_2,\sigma_1,\lambda_{2,1},\lambda_{2,2}) &= (u_2+\sigma_2 + 1)^\top u_2 + \lambda_{2,1}^\top(\sigma_1 - u_2) + \lambda_{2,2}^\top(\sigma_2 - u_1),
\end{align*}
The variational GNEs of this modified game can be obtained in closed form by solving the system of equations obtained by stacking the coupled KKT systems of player 1 and 2, with equal dual variables ($\lambda_{1,1}=\lambda_{2,1} := \lambda_1$ and $\lambda_{2,1}=\lambda_{2,2} :=\lambda_2$) \cite{auslender2000lagrangian}, i.e.,
\begin{align*}
\left[
\begin{array}{c}
\nabla_{u_1} \mc L_1(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\nabla_{\sigma_1} \mc L_1(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\nabla_{u_2} \mc \mc L_2(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\nabla_{\sigma_2} \mc \mc L_2(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\sigma_1 - u_2\\
\sigma_2 - u_1
\end{array}
\right] =
%
\left[
\begin{array}{c}
4u_1 + 2\sigma_1 +2 - \lambda_2\\
2u_1 +\lambda_1\\
2u_2 + \sigma_2 +1 - \lambda_1\\
u_2 +\lambda_2\\
\sigma_1 - u_2\\
\sigma_2 - u_1
\end{array}
\right] 
\overset{!}{
=}
%
 \mathbf{0},
\end{align*}
which yields the unique v-GNE $(u_1^*, u_2^*) = (1,-2)$. Therefore, this straightforward decomposition of the cost functions does not preserve the original equilibrium of the game.
{\hfill $\square$}
}
